---
output:
  html_document:
    df_print: paged
  pdf_document: default
  html_notebook: default
  word_document: default
  fontsize: 12pt
---


# STAT 6600: Assignment 1
## Due: September 10, 2019
## Akeem Ajede

\medskip
```{r}
Data = read.delim("Prob1InjectorPumps.txt", header = FALSE)
```

### Question 1(e)
Use R to generate a histogram of the sample distribution with the "theoretical" pdf overlaid.
\vskip

### Solution
```{r}
Raw = read.table("Prob1InjectorPumps.txt", header = F)
y = Raw[order(Raw$V1),]
hist(y, prob = T, xlab = "Plunger Relative Diameter (microns)", ylab = "count", col = "green", main = "Histogram of Plunger Relative Diameter")
m <-mean(y);std <-sqrt(var(y))
curve(dnorm(x, mean=m, sd=std), col= 2, lwd=2, add=TRUE)

```

\medskip

### Question 1(f)
Use R to generate the empirical CDF with the "theoretical" CDF overlaid.

### Solution
```{r}
plot(ecdf(y), main = "ECDF of Plunger Relative Diameter (microns)")
curve(pnorm(x, m, std), col= 2, lwd=2, add=TRUE)
```



### Question 1(g)
Use R to produce a normal QQ plot.

### Solution
```{r}
qqnorm(y)
qqline(y, col = "black")
```

\vskip
### Question 1(h)
Does the data look like it came from a normal population?l

### Solution
From the normality plot (i.e., qq - plot), the data form an approximately straight line along the normality line. Hence, it is safe to assume that the data came from a normal population.


\vskip
### Question 1(i)

Recall that since $\bar{X} \approx N(\mu_{\bar{X}},\sigma^2_{\bar{X}}) = N(\mu, \frac{\sigma^2}{n}),$ the $SE(\hat{\mu}=\frac{s}{\sqrt{n}})=$ Compute a $90\%$ confidence interval on the true mean plunger force.









