---
output:
  pdf_document: default
  fontsize: 12pt
  html_notebook: default
  html_document:
    df_print: paged
---


# STAT 6600: Assignment 1
#### Due: Tuesday, September 10, 2019
## Akeem Ajede

\medskip

Leaf Unit = 0.1

```{r}
pump_data = read.delim("Prob1InjectorPumps.txt", header = F, col.names = "Akeem")
pump_data

pump_data = as.data.frame(pump_data)
Plunger = pump_data[order(pump_data$Akeem),]
stem(Plunger, scale = 2)


```
Leaf Unit = 0.1
```{r}
hist(pump_data$Akeem, breaks = 10, freq = F, xlab = "Plunger (Microns)", ylab = "count")


lines(density(pump_data$Akeem), col = 2, lwd = 3)




```

```{r}
plot(ecdf(pump_data$Akeem), col = "blue", lwd = 2)

```

```{r}

qqnorm(pump_data$Akeem)


```


1(e). (Let $\{ Y_i \}$ be a stationary process and define $\nabla Y_t=Y_t-Y_{t-1}.$ Show that $\nabla Y_t$ has a larger variance than $Y_t$
if $\rho_y(1)< \frac{1}{2}.$
\vskip
Solution:

$var(\nabla Y_t)=var(Y_t-Y_{t-1})=var(Y_t)+var(Y_{t-1})-2cov(Y_t,Y_{t-1})$
But 
$\rho(1)=\frac{\gamma(1)}{\gamma(0)}=\frac{cov(Y_t,Y_{t-1})}{var(Y_{t-1})}$

$\implies$

$cov(Y_t,Y_{t-1})=\rho(1)var(Y_t)$

Now,
$var(\nabla Y_t)=var(Y_t-Y_{t-1})=var(Y_t)+var(Y_{t-1})-2\rho(1)var(Y_t)$

$=var(Y_t)+var(Y_{t-1})-2\rho(1)var(Y_{t-1})$
$=var(Y_t)+[1-2\rho(1)]var(Y_{t-1})$

\medskip



2. Consider the simple linear regression through the origin $y_i = i\beta/n + \epsilon_i$, with $\epsilon_i \sim N(0,1)$, $i = 1, \ldots, n$. Show that the maximum likelihood estimator is consistent for $\beta$.
\medskip

3. Let $\mathbf{X} \sim N_{n}(\boldsymbol{\mu}, \Sigma)$, where $\boldsymbol{\mu} = \mu \mathbf{1}_{n}$ and $\Sigma = \sigma^{2}[(1-\rho)I_{n} + \rho J_{n}]$, $0 \leq \rho < 1$. Show that $\bar{X}$ is distributed independently of $Q = [\sigma^{2}(1-\rho)]^{-1}\sum_{i=1}^{n}(X_{i} - \bar{X})^{2}$.
\medskip

4. Suppose $\mathbf{X} \sim N_{n}(\mathbf{0}, \sigma^{2}I_{n})$. Define $\mathbf{Y} = (Y_{1}, \ldots, Y_{n-1})'$, where $Y_{i} = \gamma X_{i} + X_{i+1}, \quad i = 1, \ldots, n-1$.
Derive the distribution of $\mathbf{Y}$.
\medskip

5. Let $\mathbf{X} = (X_1, X_2, X_3)'$ be a  random vector so that
$$ E(\mathbf{X}) = \left( \begin{array}{c}
1 \\
2 \\
3 
\end{array}
\right)
\quad
\mbox{and}
\quad Var(\mathbf{X}) = \left( \begin{array}{ccc}
5 & 2 & 3 \\
2 & 3 & 0 \\
3 & 0 & 2
\end{array}
\right) $$
Find an upper bound for $E(\|\mathbf{X}\|)$.
\medskip

6. Let $\mathbf{X}$ be as given in problem 5. Let $\mathbf{Y} = (Y_1, Y_2)'$ be a  random vector so that
$$ E(\mathbf{Y}) = \left( \begin{array}{c}
1 \\
2 
\end{array}
\right)
\quad
\mbox{and}
\quad Var(\mathbf{Y}) = \left( \begin{array}{ccc}
5 & 2 \\
2 & 3 
\end{array}
\right) $$
Find an upper bound for $E(\|\mathbf{X}\|\|\mathbf{Y}\|)$.